---
title: "Churn Analysis"
author: "Pik Shan Yiu"
date: "5/14/2019"
output: html_document
---

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(survival)
library(dplyr)
library(ggplot2)
library(scales)
library(reshape2)
library(ggcorrplot)
library(RColorBrewer)
library(randomForest)
library(caret)
library(pROC)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
raw <- read.csv('Telco_churn.csv')
```


## Exploratory Data Analysis
A EDA will provide us with more insights and allow us to establish some hypothesis to test. Let's take a quick look at the data first. 
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
str(raw)
sapply(raw, function(x) sum(is.na(x))) 
raw[is.na(raw$TotalCharges), ]
```
The quick scan of data shows that there are 11 missing values in column `TotalCharges`. An examination of these 
data report 0 in their tenure, suggesting that these are new customers who have not contributed revenues to the company yet. We can impute these missing values with 0 for now. 
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
raw$TotalCharges[is.na(raw$TotalCharges)] <- 0
```

Next, let's look at the distrbution of churned customers across different profiles. 
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(raw, aes(x = Churn))+
  geom_bar(aes(y = (..count..) / sum(..count..), fill = Churn)) +
  scale_y_continuous(labels = percent) +
  xlab("Churned Customer") +
  ylab("percentage (%)")
```

Nearly 30% of existing customers churned. Who are those who churned? 
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
raw %>% group_by(gender) %>% summarise(churnrate = sum(Churn == "Yes") / length(Churn) * 100) %>% 
  ggplot()+ geom_bar(aes(gender, churnrate, fill = gender), stat="identity") + 
  xlab("Gender") +
    ylab("Churn Rate (%)") + 
  theme(legend.position = "None")
```

No obvious gender difference in churn rate.
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
raw %>% group_by(SeniorCitizen) %>% summarise(churnrate = sum(Churn == "Yes") / length(Churn) * 100) %>% 
  ggplot()+ geom_bar(aes(SeniorCitizen, churnrate, fill = SeniorCitizen), stat="identity") + 
  xlab("Senior Citizen") +
    ylab("Churn Rate (%)") + 
  theme(legend.position = "None")
```

It sure is reasonable that senior citizen are more likely to churn, as they age they will naturally opt out of service. 
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
raw %>% group_by(Partner) %>% summarise(churnrate = sum(Churn == "Yes") / length(Churn) * 100) %>% 
  ggplot()+ geom_bar(aes(Partner, churnrate, fill = Partner), stat="identity") + 
  xlab("With Partner") +
    ylab("Churn Rate (%)") + 
  theme(legend.position = "None")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
raw %>% group_by(Dependents) %>% summarise(churnrate = sum(Churn == "Yes") / length(Churn) * 100) %>% 
  ggplot()+ geom_bar(aes(Dependents, churnrate, fill = Dependents), stat="identity") + 
  xlab("With Dependents") +
    ylab("Churn Rate (%)") + 
  theme(legend.position = "None")
```

People with partners and dependents are less likely to churn. 
<br>

How about different services?
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
corrsub <- raw %>% select(PhoneService, MultipleLines, InternetService, OnlineSecurity, OnlineBackup, DeviceProtection, TechSupport, StreamingTV, StreamingMovies, Contract, PaperlessBilling, PaymentMethod, Churn) 
corrsub <-  sapply(corrsub, as.numeric)
corrmat <-  cor(corrsub)
ggcorrplot(corrmat, hc.order = TRUE, type = "lower", outline.color = "white")
```


The correlation map shows that most extra services are correlated with lower churn, except for *paperpless billing*. In addition, longer the *contract*, shorter the churn rate. *Automatic payment* also makes customer less likely to churn. It is also worth checking if the paperless system is properly set up, or includes any inconvenient barriers. 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
par(mfrow = c(1, 3))
boxplot(tenure ~ Churn, data = raw, notch = TRUE, col = c("gold", "darkgreen"), main = "Tenure Length", xlab = "Churned")
boxplot(TotalCharges ~ Churn, data = raw, notch = TRUE, col = c("gold", "darkgreen"), main = "Total Charges", xlab = "Churned")
boxplot(MonthlyCharges ~ Churn, data = raw, notch = TRUE, col = c("gold", "darkgreen"), main = "Monthly Charges", xlab = "Churned")
```

Overall, people are less likely to churn if they are in longer tenure and paid more in total service (naturally). What's interesting, however, is the lower range of monthly charges for people who are less likely to churn, as shown in the third graph. It is hard to say for now what contributes to this difference. One possibility is that customer who enroll in month-to-month contract might pay a higher rate than those who are in longer contract, and they are more likely to churn. Let's explore this hypothesis now. 
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
boxplot(MonthlyCharges ~ Contract, data = raw, notch = TRUE, col = brewer.pal(3, "Set3"), main = "Monthly Charges", xlab = "Contract Type")
```

The graph confirms our hypothesis. Monthly contract is associated to higher monthly payment, hence related to higher churn possibility. 

## Feature Importance and Churn Prediction
We can run models to understand what are the important factors that predict customers' churn decisions, and verify with multiple models with the prediction. We can run a random forest model for this. 
```{r}
feat_subset <- raw[, -1]
feat_subset$Churn <- as.numeric(feat_subset$Churn) 
set.seed(0715)
model_rf <- randomForest(Churn ~., data = feat_subset)
rf_pred <- predict(model_rf, testing)
auc(rf_pred, testing$Churn)
model_rf$importance
```